---
title: "Exercises week 7 thursday"
author: "faho@itu.dk"
date: "2024-03-19"
output: pdf_document
---

##Exercise 1

We perform a t-test for the null hypothesis $H_0 : \mu = 10$ by means of a dataset consisting of $n = 16$ elements with sample mean 11 and sample variance 4. We use significance level 0.05.

\###(a) Should we reject the null hypothesis in favor of $H1 : \mu \neq 10?$

**Solution:**

We want to compute the $t$-statistic of the $t$-test, which is given by: $$T_{\text{obs}}=\frac{\hat{\mu}-\mu}{\text{SE}({\hat{\mu}})}=\frac{\hat{\mu}-\mu}{s/\sqrt{n}}$$

we know the sample mean and sample variance, thus we also know the sample standard deviation $s=\sqrt{s^2}=\sqrt{sqrt{4}=2$.

```{r}
T_obs <- (11-10) / (2 / sqrt(16))
print(T_obs)
```

So now we need to compute the quantile of the $t$-tidstribution $t_{1-\alpha/2,n-1}=t_{0.975,15}=2.131$

Let's also check the lower bound: $t$-tidstribution $t_{1-\alpha/2,n-1}=t_{0.025,15}=-2.131$

t distribution is symmetrical, so lower bound is not necessary really. Our observed t test statistic is 2, and the critical points are $+-2.131$ we cannot reject $H_0$.

### (b)

what if we test against $H_1 : \mu > 10$?

**Solution** So we don't need to recalculate $T_\text{obs}$, which is still equal to 2.

Instead we need to compute the $t$-quantile $t_{1-\alpha, n-1} =t_{0.95, 15} = 1.7351$.

Because $T_\text{obs}=2>1.7351$ we can now reject $H_0$ in favor of $H_1$

##Exercise 2

**Solution**

Let's first load the data

```{r}
library(UsingR)
baseball_data <- data.frame(OBP)
```

We want to compute the $t$-statistic of the $t$-test, which is given by: $$T_{\text{obs}}=\frac{\hat{\mu}-\mu}{\text{SE}({\hat{\mu}})}=\frac{\hat{\mu}-\mu}{s/\sqrt{n}}$$

```{r}
mu_hat <- mean(baseball_data$OBP)
s <- sd(baseball_data$OBP)
```

Let's do the significance test by using the ´t.test´:

```{r}
t.test(baseball_data$OBP, conf.level = 0.95, mu=0.33)
```
```{r}
qt(0.975,437)
```

Because the $p$-value is above 5% and $T_\text{obs}=-0.16846<1.965407$ we cannot reject $H_0$

